\textbf{COMP0085 Summative Assignment}

### 2022-12-27 submission.tex
add untrained and trained hyperparameter gp extrapolations
update EP and mean field questions

\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale = 0.5]{outputs/q2/a-mean}
\caption{The Posterior Mean}
\label{fig:fig2-a-mean}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale = 0.5]{outputs/q2/a-covariance}
\caption{t=The Posterior Covariance}
\label{fig:fig2-a-covariance}
\end{minipage}
\end{figure}

@@

We can see that the residuals do not perfectly conform to our prior over $e(t) \sim \mathcal{N}(0, 1)$.
The density estimation shows that a mean of zero is a reasonable prior belief however the data does not seem to exhibit unit variance.

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{outputs/q2/c-gram-matrix}
\caption{Gram Matrix}
\label{fig:c-gram-matrix}
\end{figure}

@@

\subsection*{(f)}
\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{outputs/q2/f-extrapolation}
\caption{Extrapolation of $CO_2$ concentration levels}
\label{fig:f-extrapolation}
\end{figure}

@@

\newpage
The Python code:
\lstinputlisting[language=Python]{src/solutions/q3.py}
\newpage
\section*{Question 4}

@@

\[q(\textbf{s}) = \left(\prod_{i=1}^{K} \tilde{f}_i(s_i) \right) \left(\prod_{i=1}^{K}  \prod_{j=1}^{i-1} \tilde{g}_{ij} (s_i, s_j) \right)\]

Thus, we can derive cavity distributions:

\[q_{\neg \tilde{f}_i(s_i)}(\textbf{s}) = \left(\prod_{i'=1, i'\neq i}^{K} \tilde{f}_{i'}(s_{i'}) \right) \left(\prod_{i=1}^{K}  \prod_{j=1}^{i-1} \tilde{g}_{ij} (s_i, s_j) \right)\]

@@

\[q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) = \left(\prod_{i'=1}^{K} \tilde{f}_{i'}(s_{i'}) \right) \left(\prod_{i'=1}^{K}  \prod_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1} \tilde{g}_{i'j'} (s_{i'}, s_{j'}) \right)\]

@@

\[\tilde{f}_{i}(s_{i}) = \arg \min \textbf{KL} \left[ f_{i}(s_{i}) q_{\neg \tilde{f}_i(s_i)}(\textbf{s}) \| \tilde{f}_{i}(s_{i}) q_{\neg \tilde{f}_i(s_i)}(\textbf{s}) \right]\]

We know that the factor $\log f_i(s_i)$ is a Bernoulli of the form $b_i s_i$. Because our approximation is also Bernoulli, we can simply solve for $\lambda_i$ in $\log \tilde{f}_{i}(s_{i})$:

@@

\[\tilde{g}_{ij}(s_i, s_j) = \arg \min \textbf{KL} \left[ g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \| \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \right]\]

Note that because $\tilde{g}_{ij}(s_i, s_j)$ is the product of two Bernoulli distributions, we only require the natural parameters:

@@

the mean with respect to $s_i$ and $s_j$ respectively.

@@

\[
\begin{array}{l}
 \log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \log \left(\frac{\theta_i}{1-\theta_i} \right) s_i + \log \left(\frac{\theta_j}{1-\theta_j} \right) s_j\\
\tab \tab \tab \tab \tab + \sum_{i'=1}^{K}  \log \left(\frac{\lambda_{i'}}{1-\lambda_{i'}} \right) s_{i'} \\
\tab \tab \tab \tab \tab +  \sum_{i'=1}^{K}\sum_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1}  \log \left(\frac{\theta_{i'}}{1-\theta_{i'}} \right) s_{i'} + \log \left(\frac{\theta_{j'}}{1-\theta_{j'}} \right) s_{j'}
\end{array}
\]

Only including terms with $s_i$ and $s_j$:

%\[
% \log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \left(K \log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\right) s_i +   \left(
%K \log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\right) s_j
%\]

\[
\begin{array}{l}
 \log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \left((K-1) \log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\right) s_i\\
\tab \tab \tab \tab \tab +\left(
(K-1) \log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\right) s_j
\end{array}
\]

@@

\[
\begin{array}{l}
 \log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \frac{- \mu_i^T \mu_j}{\sigma^2} s_i s_j\\
\tab \tab \tab \tab \tab + \sum_{i'=1}^{K}  \log \left(\frac{\lambda_{i'}}{1-\lambda_{i'}} \right) s_{i'} \\
\tab \tab \tab \tab \tab +  \sum_{i'=1}^{K}\sum_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1}  \log \left(\frac{\theta_{i'}}{1-\theta_{i'}} \right) s_{i'} + \log \left(\frac{\theta_{j'}}{1-\theta_{j'}} \right) s_{j'}
\end{array}
\]

Only including terms with $s_i$ and $s_j$:

\[
\begin{array}{l}
 \log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \frac{- \mu_i^T \mu_j}{\sigma^2} s_i s_j\\
\tab \tab \tab \tab \tab + \left((K-2)\log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\right) s_i \\
\tab \tab \tab \tab \tab +  \left((K-2)\log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\right) s_j
\end{array}
\]

>>>>

\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale = 0.5]{outputs/q2/a-mean}
\caption{The Posterior Mean}
\label{fig:fig2-a-mean}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale = 0.5]{outputs/q2/a-covariance}
\caption{The Posterior Covariance}
\label{fig:fig2-a-covariance}
\end{minipage}
\end{figure}

@@

We can see that the residuals do not perfectly conform to our prior over $e(t) \sim \mathcal{N}(0, 1)$.
The density estimation shows that a mean of zero is a reasonable prior belief however the data does not seem to exhibit unit variance.
Also we know it's not iid because timeseries.

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{outputs/q2/c-gram-matrix}
\caption{Covariance Matrix}
\label{fig:c-gram-matrix}
\end{figure}

@@

\subsection*{(f)}

\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale=0.5]{outputs/q2/f-extrapolation-untrained}
\caption{Without hyperparameter tuning}
\label{fig:f-extrapolation-untrained}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
\includegraphics[scale=0.5]{outputs/q2/f-extrapolation-trained}
\caption{With hyperparameter tuning}
\label{fig:f-extrapolation-trained}
\end{minipage}
\end{figure}
%
%\begin{figure}[h]
%\centering
%\includegraphics[scale=0.5]{outputs/q2/f-extrapolation-untrained}
%\caption{Extrapolation of $CO_2$ concentration levels}
%\label{fig:f-extrapolation-untrained}
%\end{figure}
%
%\begin{figure}[h]
%\centering
%\includegraphics[scale=0.5]{outputs/q2/f-extrapolation-trained}
%\caption{Extrapolation of $CO_2$ concentration levels}
%\label{fig:f-extrapolation-trained}
%\end{figure}

@@

The provided derivations for the M step of the mean parameter $\mu$:

\[\mu = \left( \langle \textbf{s}\textbf{s}^T \rangle_{q(\textbf{s})} \right)^{-1}\langle \textbf{s}\rangle_{q(\textbf{s})} \textbf{x}\]

where $\mu \in \mathbb{R}^{K \times D}$, \textbf{s} \in \mathbb{R}^{K \times N}$, and $\textbf{x} \in \mathbb{R}^{N \times D}$.

This mimics the least squares solution:

\[\hat{\beta} = (\textbf{X}\textbf{X}^T) \textbf{X} \textbf{Y}\]

for the linear regression problem $\textbf{Y} = \textbf{X}^T \beta$ where $\beta$ corresponds to the mean parameters $\mu$, the design matrix $\textbf{X}$ corresponds to the input $\textbf{s}$ and the response $Y$ corresponds to the image pixels denoted $\textbf{x}$.
This makes sense because our resulting images $\textbf{x}$ are modeled as linear combinations of features $\mu$, weighted by $\textbf{s}$.

The computational complexity of the implemented M step function can be broken down for each parameter:

\begin{itemize}
\item [$\mu$:]\begin{itemize}
\item The inversion $\text{ESS}^{-1}$ where $\text{ESS} \in \mathbb{R}^{K \times K}$ is $\mathcal{O}(K^3)$
\item The dot product $\text{ESS}^{-1} \text{ES}^T$ where $\text{ESS}^{-1} \in \mathbb{R}^{K \times K}$ and $\text{ES} \in \mathbb{R}^{N \times K}$ is  $\mathcal{O}(K^2N)$
\item The dot product $(\text{ESS}^{-1} \text{ES}^T)\textbf{x}$ where $(\text{ESS}^{-1} \text{ES}^T)\ \in \mathbb{R}^{K \times N}$ and $\textbf{x} \in \mathbb{R}^{N \times D}$ is  $\mathcal{O}(KND )$
\end{itemize}

\item [$\sigma$:] \begin{itemize}
\item The dot product $(\textbf{x}^{T} \textbf{x})$ where $\textbf{x} \in \mathbb{R}^{N \times D}$ is  $\mathcal{O}(D^2N)$
\item The dot product $\mu^{T} \mu$ where $\mu \in \mathbb{R}^{D \times K}$ is  $\mathcal{O}(K^2D)$
\item The dot product $(\mu^{T} \mu)\text{ESS}$ where $\mu^{T} \mu \in \mathbb{R}^{K \times K}$ and $\text{ESS} \in \mathbb{R}^{K \times K}$ is $\mathcal{O}(K^3)$
\end{itemize}
\item [$\pi$:] \begin{itemize}
    \item The mean operation for $\text{ES} \in \mathbb{R}^{N \times K}$ along the first dimension is $\mathcal{O}(NK)$
    \end{itemize}
\end{itemize}

Thus, the computational complexity of the M step is $\mathcal{O}(K^3+K^2N + KND + D^2N + K^2D)$ where we do not assume that any of $N$, $K$, or $D$ is large compared to the others.
\newpage
\subsection*{(d)}

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{data/images}
\caption{Images generated by randomly combined features with noise}
\label{fig:3d-images}
\end{figure}

We can plot the free energy to make sure it increases each iteration:

\subsection*{(f)}

\newpage
The Python code for mean field learning:
\lstinputlisting[language=Python]{src/models/mean_field_learning.py}

\newpage
The rest of the Python code for question 3:
\lstinputlisting[language=Python]{src/solutions/q3.py}
\newpage
\section*{Question 4}

@@

\[q(\textbf{s}) = \prod_{i'=1}^{K} \left( \tilde{f}_{i'}(s_{i'})
\prod_{j\in ne(i)}^{K} \mathcal{M}_{j \rightarrow i}
\right)
\]

where $ne(i)$ are indices of neighbouring nodes of node $i$ and $\mathcal{M}_{j \rightarrow i}$ is the message from node $j$ to $i$.

Thus, the cavity distributions:

\[q_{\neg \tilde{f}_i(s_i)}(\textbf{s}) =
\prod_{j\in ne(i)}^{K} \mathcal{M}_{j \rightarrow i}
\]

@@

\[q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) = \left( \tilde{f}_{i}(s_{i}) \prod_{k\in ne(i), k\neq j}^{K} \mathcal{M}_{k \rightarrow i}
\right)\left( \tilde{f}_{j}(s_{j}) \prod_{k\in ne(j), k\neq j}^{K} \mathcal{M}_{k \rightarrow j}
\right)
\]

@@

\[\tilde{f}_{i}(s_{i}) = \arg \min_{\tilde{f}_{i}(s_{i})} \textbf{KL} \left[ f_{i}(s_{i}) q_{\neg \tilde{f}_i(s_i)}(\textbf{s}) \| \tilde{f}_{i}(s_{i}) q_{\neg \tilde{f}_i(s_i)}(\textbf{s}) \right]\]

We know that the factor $\log f_i(s_i)$ is a Bernoulli of the form $b_i s_i$. Because our approximation for this site is also Bernoulli, we can simply solve for $\lambda_i$ in $\log \tilde{f}_{i}(s_{i})$:

@@

\[\tilde{g}_{ij}(s_i, s_j) = \arg \min_{\tilde{f}_{i}(s_{i})} \textbf{KL} \left[ g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \| \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \right]\]

Defining the natural parameters $\eta_{ij}$ and $\eta_{ji}$ for the messages:

\[\mathcal{M}_{i \rightarrow j} = \exp(\eta_{ij} s_i) = \tilde{f}_{i}(s_{i}) \prod_{k\in ne(i), k\neq j}^{K} \mathcal{M}_{k \rightarrow i} \]

\[\mathcal{M}_{j \rightarrow i}  = \exp(\eta_{ji} s_j) = \tilde{f}_{j}(s_{j}) \prod_{k\in ne(j), k\neq i}^{K} \mathcal{M}_{k \rightarrow j}\]

Note that because $\tilde{g}_{ij}(s_i, s_j)$ is the product of two Bernoulli distributions, updates to this site approximation involves updating to the natural parameters:

@@

the mean of $s_i$ and $s_j$ respectively.

@@

\[\log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \log \left(\frac{\theta_i}{1-\theta_i} \right) s_i + \log \left(\frac{\theta_j}{1-\theta_j} \right) s_j + \eta_{ij} s_i + \eta_{ji} s_j\]

Simplifying:

\[\log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \log \left(\frac{\theta_i}{1-\theta_i} + \eta_{ij} \right) s_i + \log \left(\frac{\theta_j}{1-\theta_j} + \eta_{ji} \right) s_j \]

%\[
%\begin{array}{l}
% \log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \log \left(\frac{\theta_i}{1-\theta_i} \right) s_i + \log \left(\frac{\theta_j}{1-\theta_j} \right) s_j\\
%\tab \tab \tab \tab \tab + \sum_{i'=1}^{K}  \log \left(\frac{\lambda_{i'}}{1-\lambda_{i'}} \right) s_{i'} \\
%\tab \tab \tab \tab \tab +  \sum_{i'=1}^{K}\sum_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1}  \log \left(\frac{\theta_{i'}}{1-\theta_{i'}} \right) s_{i'} + \log \left(\frac{\theta_{j'}}{1-\theta_{j'}} \right) s_{j'}
%\end{array}
%\]

%Only including terms with $s_i$ and $s_j$:
%
%\[
%\begin{array}{l}
% \log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \left((K-1) \log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\right) s_i\\
%\tab \tab \tab \tab \tab +\left(
%(K-1) \log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\right) s_j
%\end{array}
%\]

Thus, the first moments:

\[\mathbb{E}_{s_i}[\tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})] = \frac{1}{1+\exp{\left(-\left(\theta_i} + \eta_{ij}\right)}\right)}\]

\[\mathbb{E}_{s_j}[\tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})] = \frac{1}{1+\exp{\left(-\left(\theta_j + \eta_{ji}\right)}\right)}\]

@@

\[\log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto W_{ij} s_i s_j
 + \eta_{ij} s_i + \eta_{ji} s_j\]

%\[
%\begin{array}{l}
% \log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \frac{- \mu_i^T \mu_j}{\sigma^2} s_i s_j\\
%\tab \tab \tab \tab \tab + \sum_{i'=1}^{K}  \log \left(\frac{\lambda_{i'}}{1-\lambda_{i'}} \right) s_{i'} \\
%\tab \tab \tab \tab \tab +  \sum_{i'=1}^{K}\sum_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1}  \log \left(\frac{\theta_{i'}}{1-\theta_{i'}} \right) s_{i'} + \log \left(\frac{\theta_{j'}}{1-\theta_{j'}} \right) s_{j'}
%\end{array}
%\]

%Only including terms with $s_i$ and $s_j$:
%
%\[
%\begin{array}{l}
% \log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \frac{- \mu_i^T \mu_j}{\sigma^2} s_i s_j\\
%\tab \tab \tab \tab \tab + \left((K-2)\log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\right) s_i \\
%\tab \tab \tab \tab \tab +  \left((K-2)\log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\right) s_j
%\end{array}
%\]
%
%We can write this as:
%
%\[
%\log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto W_{ij} s_i s_j+ A_{s_i} s_i + B_{s_j} s_j
%\]
%
%where:
%
%\[A_{s_i} = (K-2)\log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\]
%
%and
%
%\[B_{s_j} = (K-2)\log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\]

To derive the first moment for $ g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})$ with respect to $s_i$, we first marginalise out $s_j$:

\[ \sum_{s_j \in \{0, 1\}} g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \exp\left( W_{ij} s_i + \eta_{ij} s_i + \eta_{ji}\right) +  \exp\left(\eta_{ij} s_i \right)
\]

Thus, the first moment:

\[\mathbb{E}_{s_i}[g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})] = \frac{\exp\left( W_{ij} + \eta_{ij} + \eta_{ji}\right) +  \exp\left(\eta_{ij} \right)}{[\exp\left( W_{ij} + \eta_{ij} + \eta_{ji}\right) +  \exp\left(\eta_{ij} \right)]+[\exp\left( \eta_{ji}\right) +  1]}\]

Simplifying:

\[\mathbb{E}_{s_i}[g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})] = \frac{\left( \exp\left( W_{ij} + \eta_{ji}\right) +  1 \right)\exp\left(\eta_{ij} \right)}{[\left( \exp\left( W_{ij} + \eta_{ji}\right) +  1 \right)\exp\left(\eta_{ij} \right)]+[\exp\left( \eta_{ji}\right) +  1]}\]

similarly:

\[\mathbb{E}_{s_j}[g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})] = \frac{\left( \exp\left( W_{ij} + \eta_{ij}\right) +  1 \right)\exp\left(\eta_{ji} \right)}{[\left( \exp\left( W_{ij} + \eta_{ij}\right) +  1 \right)\exp\left(\eta_{ji} \right)]+[\exp\left( \eta_{ij}\right) +  1]}\]

By setting:

\[\mathbb{E}_{s_i}[\tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})] = \mathbb{E}_{s_i}[g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})]\]

\[\mathbb{E}_{s_j}[\tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})] = \mathbb{E}_{s_j}[g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s})]\]

we can solve for the parameters of $\tilde{g}_{ij}(s_i, s_j)$ with moment matching:

\[\frac{1}{1+\exp{\left(-\left(\theta_i} + \eta_{ij}\right)}\right)}} = \frac{\left( \exp\left( W_{ij} + \eta_{ji}\right) +  1 \right)\exp\left(\eta_{ij} \right)}{[\left( \exp\left( W_{ij} + \eta_{ji}\right) +  1 \right)\exp\left(\eta_{ij} \right)]+[\exp\left( \eta_{ji}\right) +  1]}\]

\[\exp\left( \eta_{ji}\right) +  1 =\exp{\left(-\left(\theta_i + \eta_{ij}\right)}\right)}\left( \exp\left( W_{ij} + \eta_{ji}\right) +  1 \right)\exp\left(\eta_{ij} \right)\]

\[\frac{\exp\left( \eta_{ji}\right) +  1}{ \exp\left( W_{ij} + \eta_{ji}\right) +  1 } =\exp{\left(-\theta_i\right)}\]

\[\theta_i = \log \left( \frac{ 1+\exp\left( W_{ij} + \eta_{ji}\right) }{1+\exp\left( \eta_{ji}\right)} \right)\]

Similarly:

\[\theta_j = \log \left( \frac{ 1+ \exp\left( W_{ij} + \eta_{ij}\right)  }{1+\exp\left( \eta_{ij}\right)} \right)\]

Our message passing approximations:

\[\exp(\eta_{ij} s_i) = \tilde{f}_{i}(s_{i}) \prod_{k\in ne(i), k\neq j}^{K} \mathcal{M}_{k \rightarrow i} \]

\[\exp(\eta_{ji} s_j) = \tilde{f}_{j}(s_{j}) \prod_{k\in ne(j), k\neq i}^{K} \mathcal{M}_{k \rightarrow j}\]

where each message $\mathcal{M}_{j \rightarrow i}$ has a factored approximation:

\[\mathcal{M}_{k \rightarrow i} = \exp(\eta_{ki} s_k) \]

because each site $\tilde{g}_{jk}(s_j s_k)$ is approximated as a product of two messages $\mathcal{M}_{j \rightarrow k}\mathcal{M}_{k \rightarrow j}$, each a Bernoulli.

Thus, the natural parameters of the messages are updated with:

\[\eta_{ij} = b_i + \sum_{k\in ne(i), k\neq j}^{K} \eta_{ki} \]

The summation of the natural parameters of the singleton factor for node $i$ with the natural parameters of messages from all the neighbouring nodes.

This leads to a loopy BP algorithm because the nodes are fully connected (i.e. every node is the neighbour of all other nodes). Thus, we cannot simply move from one end of the graph to the other like BP for tree structured graphs.
